### 2011-02-09 paper/fitnessEvaluation.tex
More fitness, explanation of bias due to sensor area

In order to measure start and end position in the same pose,

ensure we could
start and end each run in the same position,

\acmFig{linearInterp}{1}{To ensure fair fitness measurements, the
  position of the robot was measured at the beginning and end of each
  run in the same ``ready pose'' (\figref{robot_close.jpg}).  Motion
  was interpolated linearly between this stationary pose and the
  commanded gait g(t) for one second at the beginning of the run and
  two seconds at the end, as shown above.}

\begin{tabular}{ccccc}
Ready position & a & b & c & d\\
\hline
t = 0, measure  & $t \in [0,1]$ & $t \in [1,10]$ & $t \in [10,12]$ & t = 12
\end{tabular}

\editbox{proofread. especially middle paragraph}

%\section{Experimental Evaluation}

The metric for evaluation of the designed gait was speed. To evaluate
a set of parameters, the robot was sent the parameters and instructed
to walk for a certain length of time. For each evaluation, the robot
always started and ended in the same position in order to measure true
displacement and not reward gaits the ended in a lean, since the LED
would have moved a different distance the robot. More efficient
parameters resulted in a faster gait, which translated into a longer
distance walked, measured in pixels, and a better score.

The size of the Wii remote window is approximately 175 x 120 cm. A
concern about this is that if the robot walks further than usual,
it walks outside the viewable range and thus does not count the run.
This possibility could bias the final results reported on the robot.
% probably could be better written

The only human intervention required during most learning was to
occasionally move the robot back into the viewable area of the w

>>>>

In order to measure start and end position in the same pose, we
linearly interpolated the motion of the robot between the ready
position and the commanded gait, $g(t)$.  As shown in
\figref{linearInterp}, motion for the first second was interpolated
between the ready position and $g(t)$, for the next nine seconds the
robot moved exactly as specified by $g(t)$, and for the final two
seconds its motion was interpolated smoothly back to the ready
position.

\acmFig{linearInterp}{1}{To ensure fair fitness evaluations with as
  little noise as possible, the position of the robot was measured at
  the beginning and end of each run (red circles) in the same
  \emph{ready} pose (\figref{robot_close.jpg}).  Motion was
  interpolated linearly between this stationary pose and the commanded
  gait g(t) for one second at the beginning of the run and two seconds
  at the end, as shown above.}

% finite size
% shimmy

The only human intervention required during most learning trials was
to occasionally move the robot back into the viewable area of the Wii
remote if it had walked outside.  Initially this was a rare
occurrence, as the gaits did not typically produce motion as large as
the size of the window (roughly 175 x 120 cm).  However, as gaits
improved, particularly when using HyperNEAT, the robot began to walk
out of the measurement area a non-negligible fraction of the time.
Whenever it did so, we would discard the trial and repeat it until the
gait was entirely within the window.

While this process guaranteed that we always obtained a measurement
for a given gait before proceeding, it also biased the measurements
downward.  Because the performance of the robot on a given gait varied
from trial to trial, a successful measurement was more likely to be
obtained when the gait happened to perform poorly.  This phenomenon
was negligible at first, but became significantly more pronounced when
evaluating several gaits that produced motion across nearly the entire
area.  For future studies we will likely use a 2x2 array of Wii
remotes to double the length and width of the measurement arena.

%We made one final tweak to the fitness measurement process.  HN..... 1/2
